The illusory promise of the Aligned Rank Transform

### 2025-07-18 sandbox.qmd
small correction

---
title: "The illusory promise of the Aligned Rank Transform"
subtitle: "A systematic study of rank transformations"
author: 
  - name: Theophanis Tsandilas
    orcid: 0000-0002-0158-228X
    email: theophanis.tsandilas@inria.fr
    affiliations:
      - name: Université Paris-Saclay, CNRS, Inria, LISN
        country: France
  - name: Géry Casiez
    orcid: 0000-0003-1905-815X
    email: gery.casiezuniv-lille.fr
    affiliations:
      - name: Univ. Lille, CNRS, Inria, Centrale Lille, UMR 9189 CRIStAL
        country: France
bibliography: bibliography.bib
csl: canadian-journal-of-philosophy.csl

tbl-cap-location: top
---

::: {.callout-important appearance="simple"}
## Under Review {.unnumbered}
This paper is [under review](https://www.journalovi.org/under-review.html) on the experimental track of
the [Journal of Visualization and Interaction](https://www.journalovi.org/). See the 
[reviewing process](https://github.com/journalovi/2024-tsandilas-ranktransforms/issues).
:::

@@

>>>>

```         
---
title: "The illusory promise of the Aligned Rank Transform"
author: 
  - name: Theophanis Tsandilas
    orcid: 0000-0002-0158-228X
    email: theophanis.tsandilas@inria.fr
    affiliations:
      - name: Université Paris-Saclay, CNRS, Inria, LISN
        country: France
  - name: Géry Casiez
    orcid: 0000-0003-1905-815X
    affiliations:
      - name: Univ. Lille, CNRS, Inria, Centrale Lille, UMR 9189 CRIStAL
        country: France
bibliography: bibliography.bib
csl: canadian-journal-of-philosophy.csl

tbl-cap-location: top
---
```

::: {.callout-important appearance="simple"}
## Under Review {.unnumbered}

This paper is [under review](https://www.journalovi.org/under-review.html) on the experimental track of the [Journal of Visualization and Interaction](https://www.journalovi.org/). See the [reviewing process](https://github.com/journalovi/2024-tsandilas-ranktransforms/issues).
:::

@@

Hello!

### 2025-07-18 text/40-interpretingEffects.qmd
minor correction
small correction

### Controversy about the interpretation of interactions
We further discuss interpetation issues in the broader context of generalized liner models (GLMs).

@@

**When the interpretation of effects becomes ambiguous.** Issues with interpreting interaction effects arise only when there are parallel main effects --- that is, when the coefficients of all interacting factors are non-zero. Interpretation issues may also extend to main effects when the interaction coefficient $a_{12}$ is non-zero. In such cases, we may observe a change in the marginal effect of a factor (e.g., $x_1$), even if its associated model coefficient (e.g., $a_1$) is zero.

>>>>

### Disagreement on the interpretation of effects in GLMs
We further discuss interpetation issues in the broader context of generalized linear models (GLMs).

@@

**When the interpretation of effects becomes ambiguous.** Issues with interpreting interaction effects arise only when there are parallel main effects --- that is, when the coefficients of all interacting factors are non-zero. Interpretation issues may also extend to main effects when the interaction coefficient $a_{12}$ is non-zero. In such cases, we may observe a change in the marginal effect of a factor (e.g., $x_1$), even if its associated model coefficient (e.g., $a_1$) is zero. We discuss such scenarios in more depth in the [appendix](appendix.html#removable-main).
